---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "100%"
)
```

# wbdataset

<!-- badges: start -->
[![Lifecycle: experimental](https://img.shields.io/badge/lifecycle-experimental-orange.svg)](https://lifecycle.r-lib.org/articles/stages.html#experimental)
[![CRAN status](https://www.r-pkg.org/badges/version/wbdataset)](https://CRAN.R-project.org/package=wbdataset)
[![DOI](https://zenodo.org/badge/DOI/10.5281/zenodo.13972192.svg)]
[![Codecov test coverage](https://codecov.io/gh/dataobservatory-eu/wbdataset/graph/badge.svg)](https://app.codecov.io/gh/dataobservatory-eu/wbdataset)
<!-- badges: end -->

The goal of `wbdataset` is to create tidy datasets from Wikidata or a Wikibase instance. The `wbdataset` package is an extension of the `dataset`, which in turn is an R package that helps to exchange, publish and combine datasets more easily by improving their semantics. The `wbdataset` extends the usability of dataset by connecting the Wikibase API with the R statistical environment.

## Installation

You can install the development version of wbdataset from [GitHub](https://github.com/) with:

``` r
# install.packages("devtools")
devtools::install_github("antaldaniel/dataset")
devtools::install_github("dataobservatory-eu/wbdataset")
```

```{r}
library(wbdataset)
```

## Example

This is a basic example which shows you how to solve a common problem:

```{r example}
library(wbdataset)
get_item(qid=c("Q228", "Q347"), 
         language=c("en", "nl"), 
         creator=person("Jane Doe"), 
         title="Small Countries")
```



