---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "100%"
)
```

# wbdataset

<!-- badges: start -->
[![Lifecycle: experimental](https://img.shields.io/badge/lifecycle-experimental-orange.svg)](https://lifecycle.r-lib.org/articles/stages.html#experimental)
[![CRAN status](https://www.r-pkg.org/badges/version/wbdataset)](https://CRAN.R-project.org/package=wbdataset)
[![devel-version](https://img.shields.io/badge/devel%20version-0.1.1014-blue.svg)](https://github.com/dataobservatory-eu/wbdataset)
[![DOI](https://zenodo.org/badge/DOI/10.5281/zenodo.13972192.svg)]
[![Codecov test coverage](https://codecov.io/gh/dataobservatory-eu/wbdataset/graph/badge.svg)](https://app.codecov.io/gh/dataobservatory-eu/wbdataset)
<!-- badges: end -->

The goal of `wbdataset` is to create tidy datasets from Wikidata or a Wikibase instance. The `wbdataset` package is an extension of the `dataset`, which in turn is an R package that helps to exchange, publish and combine datasets more easily by improving their semantics. The `wbdataset` extends the usability of dataset by connecting the Wikibase API with the R statistical environment.

## Installation

You can install the development version of wbdataset from [GitHub](https://github.com/) with:

``` r
# install.packages("devtools")
devtools::install_github("antaldaniel/dataset")
devtools::install_github("dataobservatory-eu/wbdataset")
```

```{r}
library(wbdataset)
```

## Example

This is a basic example which shows you how to solve a common problem:

```{r example}
library(wbdataset)
small_countries_dataset <- get_wikidata_item(
  qid = c("Q228", "Q347"),
  language = c("en", "nl"),
  data_curator = person("Jane Doe"),
  title = "Small Countries"
)

small_countries_dataset
```
The *wbdataset* package has three tutorial articles:

- [Importing from Wikidata or Wikibase](https://wbdataset.dataobservatory.eu/articles/wikidata-import.html) shows you how to import data from Wikidata.
-  [Authenticated MediaWiki API Access](https://wbdataset.dataobservatory.eu/articles/Wikibase_API.html) shows how you log in to a Wikibase instance. If you use `wbdataset` to retrieve information only from Wikidata, you do not need this.
- [Getting Started To Create a Graph on Wikibase](https://wbdataset.dataobservatory.eu/articles/start.html) shows how to add data to a Wikibase graph database.


