---
title: "Importing from Wikidata or Wikibase"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Importing from Wikidata or Wikibase}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r setupvignette, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

The _wbdataset_ package is an extension of the dataset, which in turn is an R package that helps to exchange, publish and combine datasets more easily by improving their semantics. The wbdataset extends the usability of dataset by connecting the Wikibase API with the R statistical environment.

Let us initialize a dataset from Wikidata.

```{r setup}
library(dataset)
library(wbdataset)
data("wikidata_countries_df")
```

You do not need to use the login functions `get_csrf` and `get_csrf_token` to work with Wikidata, however, you need to provide login credentials with those functions to a privately managed Wikibase instance. In this example, we retrieve data from Wikidata without password protection. The same scripts would retrieve data from a private Wikibase with providing the optional CSRF token to the functions.

## Retrieving items

Items will serve as your observations or data subjects.  We will collect their attributes (usually constants, like the area size of a country) and variables (like their population.)

We will start with four countries, of course, you could read in a longer list from a file:

```{r country-select}
# Select the following country profiles from Wikidata:
wikidata_countries <- c(
  "http://www.wikidata.org/entity/Q756617", "http://www.wikidata.org/entity/Q347",
  "http://www.wikidata.org/entity/Q3908",   "http://www.wikidata.org/entity/Q1246"
)
```

And download the main identifiers of these data subject, i.e., the countries:

- `QID`:
- `label`:
- `description`:

```{r get-data, eval=FALSE}
# Retrieve their labels into a dataset called 'European countries':
wikidata_countries_df <- get_wikidata_item(
  qid = wikidata_countries,
  language = "en",
  title = "European countries",
  creator = person("Daniel", "Antal")
)
```

The resulting dataset:

```{r print}
print(wikidata_countries_df)
```

The provenance and the definition of the key `qid` column is well described in the attributes. 

### Adding attributes and variables

Now let us add further columns, making sure that we include the precise definition of each of the variables.

We add properties of these countries, or attributes/variables of these countries by reading statements about those variables or attributes. Our `left_join_column` tries to retrieve the attribute, for example, `ISO 3166-1 alpha-2 code (P297)` for every country identified by a QID, starting with `Q756617` as defined above.

```{r addcols, eval=FALSE}
ds <- wikidata_countries_df %>%
  left_join_column(
    label = "ISO 3166-1 alpha-2 code",
    property = "P297"
  ) %>%
  left_join_column(
    property = "P1566"
  ) 
```

You can set the workflow to SILENT to avoid the progress report above, i.e.,  
`Left join claims: m/n: Qxxx Pyyy`.

```{r addcols-silent, eval=FALSE}
ds <- wikidata_countries_df %>%
  left_join_column(
    label = "ISO 3166-1 alpha-2 code",
    property = "P297", silent = TRUE
  ) %>%
  left_join_column(
    property = "P1566",
    label = "Geonames ID",
    namespace = "https://www.geonames.org/", silent = TRUE
  ) %>%
  left_join_column(
    label = "different from Wikipedia item",
    property = "P1889", silent = TRUE
  )
```


### Saving the data 

Saving the data in rds or rda files will retain the rich metadata:

```{r save, eval=FALSE}
saveRDS(ds, file = tempfile())
```

